General constraints on magic distillation protocols from statistical mechanics

### 2021-04-28 magic_paper.tex
Added assumptions, statement of new theorem, altered first elbow condition to match assumptions. Next: Add proof, push technical details to appendix

\begin{document}
% Write bounds as 1 + ln{}/{} or 1 - (-ln{})/{} ?
% Sec: General resources
% Sec: Lower bounds
% App: D clean-up + d--> r ?

% Figs consistency (esp. figs 3,4,5 and 1,6) + Figure 0 ?

% Eq. numbering in last line always
% UK vs US english
% Spacing (fig positioning, sections, paragraphs, QED symbols)

@@

Central to the analysis is the notion of a Lorenz curve of a vector $\w \in \mathbb{R}^n$ relative to some other vector $\r \in \mathbb{R}^n$. Given a vector $\w$ we define $\w^\downarrow$ to be the re-arrangement of the components of $\w$ into decreasing order. Given two $n$--component vectors $\w$ and $\r$, we first define $\widetilde{\w}$, where $\widetilde{w}_i := w_i/r_i$, as the vector of component-wise ratios between $\w$ and $\r$.
We can now define the \emph{Lorenz curve} of $\w$ relative to $\r$, denoted $L_{\w|\r}(x)$, as the piece-wise linear function that passes through $(0,0)$ and the $n$ points
\begin{equation}
\label{eq:lc}
        (x_k,L_{\w|\r}(x_k)) =\left( \frac{1}{R}\sum_{i=1}^k r_{\pi(i)}, \sum_{i=1}^k w_{\pi(i)} \right),
\end{equation}
where $R:= \sum_{i=1}^n r_i$ and $\pi$ is the permutation on $n$ objects mapping $\widetilde{\w}$ to $\widetilde{\w}^{\downarrow}$. The form of this requires that $\r$ has no zero components, which we shall assume without loss of generality as in any physical situation the rank of a quantum state is not an operationally meaningful quantity.
We also observe that the curve contains non-differentiable points, called \emph{elbows} at locations where $\widetilde{\bmw}$ changes value.

@@

\begin{figure}
    \centering
    \includegraphics[scale=0.35]{figs/lc_strange.pdf}
    \caption{\textbf{A `heretical' family of Lorenz curves.} Traditionally, Lorenz curves for distributions are monotone increasing cumulant functions that reach a maximum value of $1$. In contrast, Lorenz curves for magic states break through the value of $L(x)=1$ due to the presence of negativity in the associated quasi-probability distributions. The above family of curves correspond to multiple copies of noisy Strange states $\rho:=\rho_{\rm{S}}(\epsilon)^{\otimes n}$ (\cref{eq:noisysn}) for $n=2,4,6$ within a particular stabilizer fragment ($\sigma = \id/3$). Solid lines represent pure Strange states, while dashed lines represent noisy Strange states with depolarising noise ($\epsilon = 0.1$).
    }
    \label{fig:lcs}
\end{figure}

@@

More precisely, since the components of $\r$ are strictly positive, there always exists an $\epsilon >0$ such that $\w_\epsilon := \epsilon \w + (1-\epsilon) \r$ is a genuine probability distribution. We can see that $(\w , \r) \succ (\w', \r')$ if and only if $(\w_\epsilon , \r) \succ (\w_\epsilon', \r')$, where we choose $\epsilon$ so that both $\w_\epsilon$ and $\w_\epsilon'$ are probability distributions. This equivalence holds because there exists a stochastic map $A$ such that $A \w = \w'$ and $A \r = \r'$ if and only if 
\begin{equation}
A[\epsilon \w + (1-\epsilon) \r] = \epsilon \w' + (1-\epsilon) \r'\mbox{ and } A\r = \r'.
\end{equation}
As mentioned, relative majorization between pairs of distributions occurs if and only if the associated Lorenz curves never cross. Therefore, we can express this condition in terms of the Lorenz curves of the quasi-probability distributions $\w$ and $\w'$ relative to $\r$ and $\r'$ respectively.

@@

\begin{lemma} Relative majorization in any fragment implies the monotonicity of sum-negativity. 
\end{lemma}
\begin{proof}
	The sum-negativity of a magic state $\rho$ can be written as $sn(\rho) := \frac{1}{2} (\sum_{\x} |W_\rho(\x) | - 1)$.
We make use of a variant form of relative majorization, (see~\cref{prop:rmajor} in~\cref{app:major}), stating that $\p \succ_{\r} \q$ if and only if
	\begin{equation}
\sum_k | p_k - r_k t | \geq \sum_k | q_k - r_k t |,
\end{equation}
for all $t\in \mathbb{R}$. Choosing $t=0$ we get the single condition that $\sum_k |p_k| \ge |\sum_k |q_k|$, independent of the choice of $\r$. Applying this to the Wigner quasi-distributions of two quantum states immediately gives the result.
\end{proof}
As a corollary, this result applies to mana too, since mana is an increasing function of sum-negativity.

@@

For our distillation protocols we need to choose a canonical pure magic state as a target. For this we shall choose the Strange state $|S\>$ given as
\begin{equation}
|S\> := \frac{1}{\sqrt{2}} (|1\> + |2\>),
\end{equation}
in the computational basis. This qutrit magic state is exceptionally symmetric and simple which benefits our analysis. Its distribution $W_S(\x)$ has a single negative value of $-1/3$ at $\x =(0,0)$ and the positive value $1/6$ at all other points. We can define the $\epsilon$--noisy Strange state as
\begin{equation}\label{eq:noisysn}
    \rho_{\rm{S}}(\epsilon)\coloneqq (1 - \epsilon) \ket{\rm{S}}\bra{\rm{S}} + \epsilon \frac{1}{3}\id,
\end{equation}
where $\epsilon$ is the depolarization noise parameter. Moreover, utilising the symmetries of the Strange state allow for any magic state $\rho$ to be processed via Clifford operations~\cite{cit:prakash,cit:prakash2} and put into this canonical form for some $\epsilon \ge 0$.

@@

@@

The case of correlations between output subsystems can also be considered, but would lead to a more complex analysis within our majorization framework of Wigner distributions. In this direction, we highlight recent work in majorization in which stochastic independence and correlations are analysed. In particular it has been shown that stochastic independence (no correlations) of independent distributions can be viewed as a resource in an extension of catalytic majorization, and leads to a single-shot operational interpretation of the Shannon entropy. \nick{CITE}

The action of the protocol on the reference states therefore transforms $\tau \rightarrow \sigma$. It turns out our distillation bounds depend on the free energy of the input and output states, so for a state $\tau = e^{-\beta H}/\Z$ we express the associated free energy $F$ by
\begin{equation}
	F \coloneqq \tr[H \tau] - \beta^{-1}S(\tau) = -\beta^{-1}\log{\Z}.
\end{equation}
The output state $\sigma$ is generally a non-equilibrium state for the system. However, the distillation bound takes a particularly simple form if we associate a similar free energy value to the output state. We do this by considering the change in Hamiltonian $H \rightarrow H'$ such that equilibrium is restored at the reference temperature $T$. This implies that 
\begin{equation}
\sigma = \frac{e^{-\beta H'}}{\Z'},
\end{equation}
which has free energy $F' = -\beta^{-1} \log \Z'$. Therefore the action of $\E$ sending $\tau$ to $\sigma$ is viewed in terms of $\Delta H := H'-H$, which is the Hamiltonian change that would be needed to adiabatically restore equilibrium at temperature $T$.

 Given this account of the physical ingredients of a distillation protocol we now state the following result for the case of qutrits.
\begin{theorem}\label{thm:free-energy}
	Consider a magic distillation protocol transforming noisy Strange state
	\begin{equation*}
		\rho_S(\epsilon)^{\otimes n} \longrightarrow \E(\rho_S(\epsilon)^{\otimes n})=\rho_S(\epsilon')^{\otimes m} 
	\end{equation*}
	with $n, m \gg 1$.

@@

As a special case of the result, suppose that $H=H'$, implying that the protocol leaves the equilibrium state unchanged, hence it corresponds to a Gibbs-preserving map~\cite{faist_2015}. In the more general case, $\Delta F := F' - F \ne 0$ and we have a free energy difference between the initial and final states, with respect to the relevant Hamiltonians. This is similar to the free energy differences that occur in fluctuation relations, where the final state of the system can be arbitrarily far from equilibrium, yet a free energy term for the final system still arises.

The above bounds mainly demonstrate how one can use relative majorization of quasi-distributions to incorporate additional information into distillation bounds. However to unpack the results more it is useful to consider the case in which $H=H'$, and look at $\epsilon' \rightarrow 0$. For this we obtain~\cref{fig:rate_contour} which gives a contour plot of $R$ as a function of inverse temperature $\beta$ and depolarising noise level $\epsilon$ for the initial noisy magic state.
\begin{figure}[t!]
    \centering
    \includegraphics[scale=0.45]{figs/rate_scatter.pdf}
    \caption{\textbf{Temperature-dependent bounds for magic distillation.}
  Shown is a contour plot of  $R(\epsilon, \beta)$ for $H=H'$, where $\beta$ is an inverse temperature and $\epsilon$ is the depolarising noise of the magic state. The relevant data from the spectrum of the qutrit operator $H$ enters in via $\epsilon_\star(\beta)$ and $\beta_\star$.  The vertical dashed line is the `Landauer-like' temperature threshold $\beta_\star$ and the diagonal dashed curve corresponds to the noise threshold $\epsilon_\star (\beta)$ for $\beta \leq \beta_\star$. The $\beta =0 $ line corresponds to the regime in which the magic protocol gives a unital channel.
    }
    \label{fig:rate_contour}
\end{figure}
\begin{proof}
To establish the distillation bounds we consider the distillation protocol that gives
\begin{equation}
	\E( \rho_S(\epsilon)^{\otimes n}) = \rho_S(\epsilon')^{\otimes m},
\end{equation}
for the magic state. We then consider how the protocol transforms the reference equilibrium state as
\begin{equation}
	\E( \tau^{\otimes n}) = \sigma^{\otimes m}.
\end{equation}
Since $\tau$ and $\sigma$ are assumed to be full rank stabilizer states they have a strictly positive Wigner distribution, while, in contrast, the input and output magic states will generally have quasi-probability distributions for their Wigner functions. For any such protocol we therefore have that
\begin{equation}
	(W_{\rho_n} (x), W_{\tau_n}(x) ) \succ  (W_{\rho'_m} (x), W_{\sigma_m}(x) ),
\end{equation}
where for simplicity we write $\rho_n := \rho_S(\epsilon)^{\otimes n}$, $\rho'_m:= \rho_S(\epsilon')^{\otimes m}$ and $\tau_n:=\tau^{\otimes m}$ and $\sigma_m:=\sigma^{\otimes m}$. We then have that
\begin{equation}
	L_{\rho_n |\tau_n}(x) \ge L_{\rho'_m |\sigma_m}(x) \mbox{ for all } x.
\end{equation}
We must therefore compute the Lorenz curve data for $\rho_n$ relative to $\tau_n$, and compare with the Lorenz curve of the output state $\rho_m'$ relative to $\sigma_m$.

We define $W_{\rho | \tau}(\x) := W_\rho(\x)/W_\tau(\x)$, which is always well-defined since $\tau$ is full-rank. Now we show in~\cref{app:wigner} that
\begin{equation}
W_{\rho_1\otimes \rho_2 | \tau_1 \otimes \tau_2} (\x_1 \oplus \x_2) = W_{\rho_1 | \tau_1}(\x_1)W_{\rho_2 | \tau_2}(\x_2),
\end{equation}
and also,
\begin{equation}
W_{\rho_1\otimes \rho_2} (\x_1 \oplus \x_2) = W_{\rho_1}(\x_1)W_{\rho_2}(\x_2),
\end{equation}
for any states $\rho_1, \rho_2$ and any full-rank stabilizer states $\tau_1, \tau_2$. \ddd{[Notation clashing here. Frame result outside as its own lemma -- we gotta check everything.]} \nick{Will frame in~\cref{app:wigner} and fix notation in the next commit}
Therefore we have that
\begin{align}
W_{\rho_n |\tau_n} (\x) &= \prod_{i=1}^n W_{\rho|\tau}(\x_i)\\
W_{\rho_n } (\x) &= \prod_{i=1}^n W_\rho(\x_i)
\end{align}
where $\x = \oplus_{i=1}^n \x_i$ is the phase space point for the full system in terms of those of the individual subsystems.

@@

Generally this is complex, but in order to obtain distillation bounds it is sufficient to determine the location of the first elbow $(x_0, L_0)$ of $L_{\rho_n|\tau_n}(\x)$. To do so, we compute the largest component 
\begin{equation}
	w_{\rm max} := \max_{i,j,k} [\w(\rho_n | \tau_n)]_{ijk},
\end{equation}
and determine the indices at which this occurs.
Putting in the values we obtain
\begin{align}
	&(3\Z)^{-n}w_{\rm max} = \nonumber\\
	&\max\limits_{i,j,k}\Big\{ (-v)^{n-\alpha} u^{\alpha} e^{\beta (n-\alpha)E_s} e^{\beta ( i E_0 + j E_1 + k E_2 )} \Big\}, \label{eq:max_slope}
\end{align}
where $0 \leq i,j,k \leq n$ and $\alpha \coloneqq i+j+k \leq n$.
Now for $0 \leq \epsilon \leq 3/7$, we have $v \geq u$. Since we assume that $n$ is even, we need the sum $\alpha = i+j+k$ to be even too, so that the expression is positive.

@@

The rescaled distribution is then $\tilde{\w}:= (w_i/t_i)$, and we define $v_0 := v/t_0$ and $u_i := u/t_i$ for $i\ne 0$. We now need to consider the $n$--copy version of this, denoted $\tilde{\w}^{\otimes n}$, and determine how the largest component arises. A general component of $\tilde{\w}^{\otimes n}$ is given by,
\begin{equation}
\tilde{w}_{n,\alpha, i_1, \dots, i_8} = (-v_0)^{i_0} u_1^{i_1} u_2^{i_2} \cdots u_8^{i_8},
\end{equation}
with the assumption that firstly $n$ is always even, and the indices obeying the following conditions:
\begin{align}
0 &\le i_j \le n \mbox{ for each }j=0,1,\dots ,8 \nonumber \\
i_0+ i_1 &+ i_2 + \dots i_8 =n.
\end{align}
Now the largest entry must be positive, and so we require that $\alpha$ is an even number, and so we consider
\begin{equation}
\tilde{w}_{n,i_0, i_1, \dots, i_8} = v_0^{i_0} u_1^{i_1} u_2^{i_2} \cdots u_8^{i_8},
\end{equation}
over the above integers with $i_0 \in \{0,2,4, \dots, n\}$. Let $a = \max \{v_0, u_1, \dots ,u_8\}$, then clearly the maximum value attained ($\tilde{w}_{\rm max}:= \max_{i_0,i_1,\dots , i_8}(\tilde{w}_{n,i_0, i_1, \dots, i_8})$ ) is given by 
\begin{equation}
\tilde{w}_{\rm max} = a^n.
\end{equation}
If the numbers $\{v_0, u_1, u_2, \dots, u_8\}$ are distinct then this maximum value occurs at a unique $i_k=n$ for some $k \in \{0, 1, \dots, 8\}$. Let's consider this case first.

>>>>

\begin{document}
% Write bounds as 1 + ln{}/{} or 1 - (-ln{})/{} ?
% R: rate vs bound ?
% Appendices: clean-up (d --> r)

% Figs check(fig.0, move to app, consistency)
% Eq. numbering in last line always
% UK vs US english
% Spacing (fig positioning, sections, paragraphs, QED symbols)

@@

Central to the analysis is the notion of a Lorenz curve of a vector $\w \in \mathbb{R}^n$ relative to some other vector $\r \in \mathbb{R}^n$. Given a vector $\w$ we define $\w^\downarrow$ to be the re-arrangement of the components of $\w$ into decreasing order. Given two $n$--component vectors $\w$ and $\r$, we first define $\widetilde{\w}$, where $\widetilde{w}_i \coloneqq w_i/r_i$, as the vector of component-wise ratios between $\w$ and $\r$.
We can now define the \emph{Lorenz curve} of $\w$ relative to $\r$, denoted $L_{\w|\r}(x)$, as the piece-wise linear function that passes through $(0,0)$ and the $n$ points
\begin{equation}
\label{eq:lc}
        (x_k,L_{\w|\r}(x_k)) =\left( \frac{1}{R}\sum_{i=1}^k r_{\pi(i)}, \sum_{i=1}^k w_{\pi(i)} \right),
\end{equation}
where $R\coloneqq \sum_{i=1}^n r_i$ and $\pi$ is the permutation on $n$ objects mapping $\widetilde{\w}$ to $\widetilde{\w}^{\downarrow}$. The form of this requires that $\r$ has no zero components, which we shall assume without loss of generality as in any physical situation the rank of a quantum state is not an operationally meaningful quantity.
We also observe that the curve contains non-differentiable points, called \emph{elbows} at locations where $\widetilde{\bmw}$ changes value.

@@

\begin{figure}
    \centering
    \includegraphics[scale=0.35]{figs/lc_strange.pdf}
    \caption{\textbf{A `heretical' family of Lorenz curves.} Traditionally, Lorenz curves for distributions are monotone increasing cumulant functions that reach a maximum value of $1$. In contrast, Lorenz curves for magic states break through the value of $L(x)=1$ due to the presence of negativity in the associated quasi-probability distributions. The above family of curves correspond to multiple copies of noisy Strange states $\rho\coloneqq\rho_{\rm{S}}(\epsilon)^{\otimes n}$ (\cref{eq:noisysn}) for $n=2,4,6$ within a particular stabilizer fragment ($\sigma \coloneqq \id/3$). Solid lines represent pure Strange states, while dashed lines represent noisy Strange states with depolarising noise ($\epsilon = 0.1$).
    }
    \label{fig:lcs}
\end{figure}

@@

More precisely, since the components of $\r$ are strictly positive, there always exists an $\epsilon >0$ such that $\w_\epsilon \coloneqq \epsilon \w + (1-\epsilon) \r$ is a genuine probability distribution. We can see that $(\w , \r) \succ (\w', \r')$ if and only if $(\w_\epsilon , \r) \succ (\w_\epsilon', \r')$, where we choose $\epsilon$ so that both $\w_\epsilon$ and $\w_\epsilon'$ are probability distributions. This equivalence holds because there exists a stochastic map $A$ such that $A \w = \w'$ and $A \r = \r'$ if and only if 
\begin{equation}
A[\epsilon \w + (1-\epsilon) \r] = \epsilon \w' + (1-\epsilon) \r'\mbox{ and } A\r = \r'.
\end{equation}
As mentioned, relative majorization between pairs of distributions occurs if and only if the associated Lorenz curves never cross. Therefore, we can express this condition in terms of the Lorenz curves of the quasi-probability distributions $\w$ and $\w'$ relative to $\r$ and $\r'$ respectively.

@@

\begin{lemma} Relative majorization in any fragment implies the monotonicity of sum-negativity. 
\end{lemma}
\begin{proof}
	The sum-negativity of a magic state $\rho$ can be written as $sn(\rho) \coloneqq \frac{1}{2} (\sum_{\x} |W_\rho(\x) | - 1)$.
We make use of a variant form of relative majorization, (see~\cref{prop:rmajor} in~\cref{app:major}), stating that $\p \succ_{\r} \q$ if and only if
	\begin{equation}
\sum_k | p_k - r_k t | \geq \sum_k | q_k - r_k t |,
\end{equation}
for all $t\in \mathbb{R}$. Choosing $t=0$ we get the single condition that $\sum_k |p_k| \ge |\sum_k |q_k|$, independent of the choice of $\r$. Applying this to the Wigner quasi-distributions of two quantum states immediately gives the result.
\end{proof}
As a corollary, this result applies to mana too, since mana is an increasing function of sum-negativity.

@@

For our distillation protocols we need to choose a canonical pure magic state as a target. For this we shall choose the Strange state $|S\>$ given as
\begin{equation}
|S\> \coloneqq \frac{1}{\sqrt{2}} (|1\> + |2\>),
\end{equation}
in the computational basis. This qutrit magic state is exceptionally symmetric and simple which benefits our analysis. Its distribution $W_S(\x)$ has a single negative value of $-1/3$ at $\x =(0,0)$ and the positive value $1/6$ at all other points. We can define the $\epsilon$--noisy Strange state as
\begin{equation}\label{eq:noisysn}
    \rho_{\rm{S}}(\epsilon)\coloneqq (1 - \epsilon) \ket{\rm{S}}\bra{\rm{S}} + \epsilon \frac{1}{3}\id,
\end{equation}
where $\epsilon$ is the depolarization noise parameter. Moreover, utilising the symmetries of the Strange state allow for any magic state $\rho$ to be processed via Clifford operations~\cite{cit:prakash,cit:prakash2} and put into this canonical form for some $\epsilon \ge 0$.

@@

As it is important for the statement of our result, we hasten to note that group averaging of the output system and, indeed, any permutation of the output subsystems can be performed via Clifford unitaries, for example $\rm{SWAP}$ gates, which do not affect the magic output of the protocol, so we consider distillation protocols \emph{modulo Clifford operations}.

@@

The case of correlations between output subsystems can also be considered, but would lead to a more complex analysis within our majorization framework of Wigner distributions. In this direction, we highlight recent work in majorization in which stochastic independence and correlations are analysed. In particular it has been shown that stochastic independence (no correlations) of independent distributions can be viewed as a resource in an extension of catalytic majorization, and leads to a single-shot operational interpretation of the Shannon entropy. \nick{cite David?}

The action of the protocol on the reference state therefore transforms $\tau^{\otimes n} \rightarrow \E(\tau^{\otimes n}) = \sigma^{\otimes m}$. 
The output state $\sigma$ is a stabilizer state, provided that $\tau$ is as well and that $\E$ is a free operation, but it is generally a non-equilibrium state for the system. However, the distillation bounds that we derive take a particularly simple form if we associate a fiducial Hamiltonian $H'$ to the output state by considering the change $H \rightarrow H'$ such that equilibrium is restored at the reference temperature $T$. 
This allows us to express the state as $\sigma = e^{-\beta H'}/\Z'$.

Our distillation bounds depend on two properties of the Hamiltonians $H, H'$.
The first property is the free energy of the states at the reference temperature, so for a state $\tau = e^{-\beta H}/\Z$ we express the associated free energy $F$ as
\begin{equation}
	F \coloneqq \tr[H \tau] - \beta^{-1}S(\tau) = -\beta^{-1}\log{\Z}.
\end{equation}
The output state $\sigma$ then has free energy $F' = -\beta^{-1} \log \Z'$.

The bounds also depend on the overlaps between the Hamiltonian eigenbases and the computational basis $\{\ket{k}\}$, where $k \in \mathbb{Z}_d$.
For a reference state $\tau$ with associated Hamiltonian $H$ denote the energy eigenbasis by $\{\ket{\phi_k}\}$ and define the \emph{reference partition function $\zeta$} as
\begin{equation}
	\zeta \coloneqq \sum_{k \in \mathbb{Z}_d} \alpha_k e^{-\beta E_k},
\end{equation}
where the overlap coefficients are
\begin{equation}
	\alpha_k \coloneqq \sum_{\mathbb{Z}_d} \braket{\phi_k}{-r}\braket{r}{\phi_k}.
\end{equation}
Coefficients $\alpha_k$ are not necessarily positive, which is the case when the Hamiltonian $H$ does not admit a stabilizer decomposition. 
However, it is simple to see that $\zeta$ is always positive. \nick{because it is proportional to $W_{\tau}(0,0) > 0$ for full-rank, stabilizer $\tau$ - this argument is a bit cyclic}
We can then naturally associate to the reference state $\tau$ a new form of free energy which is specific to the protocol, which we call the \emph{reference free energy $\Phi$} and define via
\begin{equation}
	\Phi \coloneqq -\beta^{-1} \log \zeta.
\end{equation}
For the output reference state $\sigma$ with associated Hamiltonian $H'$, we can denote the energy eigenbasis by $\{\ket{\phi'_k}\}$ and similarly define the reference partition function $\zeta'$ and reference free energy $\Phi'$.

Therefore, the action of $\E$ sending $\tau^{\otimes n}$ to $\sigma^{\otimes m}$ is viewed in terms of $\Delta H \coloneqq H'-H$, which is the Hamiltonian change that would be needed to adiabatically restore equilibrium at temperature $T$.
Given this account of the physical ingredients of a distillation protocol, we now state the following result for the case of qutrits.
\begin{theorem}\label{thm:free-energy}
	Consider a magic distillation protocol that transforms the noisy Strange state
	\begin{equation*}
		\rho_S(\epsilon)^{\otimes n} \longrightarrow \E(\rho_S(\epsilon)^{\otimes n})=\rho_S(\epsilon')^{\otimes m} 
	\end{equation*}
modulo Clifford operations, with even number of copies $n, m$ and noise levels satisfying $\epsilon' \leq \epsilon \leq 3/7$.

Let $T =(k\beta)^{-1}$ be any characteristic temperature for the physical system and let $H$ be the Hamiltonian of each qutrit subsystem.
Assume that $n,m \gg 1$, so that the channel $\E$ generates negligible correlations on some product equilibrium state 
\begin{equation}
	\tau^{\otimes n} \longrightarrow \E(\tau^{\otimes n}) = \sigma^{\otimes m}
\end{equation} 
for some state $\sigma = e^{-\beta H'}/\Z'$.

Then, the distillation rate $R = m/n$ of the magic protocol is bounded as:
\begin{equation}\label{eq:rate_bounds_proof}
	R \leq \dfrac{\ln{\big( 1-\frac{4}{3}\epsilon \big)} + \beta (\Phi - F)}{\ln{\big( 1-\frac{4}{3}\epsilon' \big)} + \beta (\Phi' - F')},
\end{equation}
where $F, \Phi$ are the free energy and reference free energy defined for state $\tau$, while $F', \Phi'$ are defined for state $\sigma$.
\end{theorem}
\begin{proof}
	\nick{Start}
\end{proof}

\begin{theorem}\label{thm:free-energy}
	Consider a magic distillation protocol transforming noisy Strange state
	\begin{equation*}
		\rho_S(\epsilon)^{\otimes n} \longrightarrow \E(\rho_S(\epsilon)^{\otimes n})=\rho_S(\epsilon')^{\otimes m} 
	\end{equation*}
	with $n, m \gg 1$.

@@

As a special case of the result, suppose that $H=H'$, implying that the protocol leaves the equilibrium state unchanged, hence it corresponds to a Gibbs-preserving map~\cite{faist_2015}. In the more general case, $\Delta F \coloneqq F' - F \ne 0$ and we have a free energy difference between the initial and final states, with respect to the relevant Hamiltonians. This is similar to the free energy differences that occur in fluctuation relations, where the final state of the system can be arbitrarily far from equilibrium, yet a free energy term for the final system still arises.

The above bounds mainly demonstrate how one can use relative majorization of quasi-distributions to incorporate additional information into distillation bounds. However to unpack the results more it is useful to consider the case in which $H=H'$, and look at $\epsilon' \rightarrow 0$. For this we obtain~\cref{fig:rate_contour} which gives a contour plot of $R$ as a function of inverse temperature $\beta$ and depolarising noise level $\epsilon$ for the initial noisy magic state.
\begin{figure}[t!]
    \centering
    \includegraphics[scale=0.45]{figs/rate_scatter.pdf}
    \caption{\textbf{Temperature-dependent bounds for magic distillation.}
  Shown is a contour plot of  $R(\epsilon, \beta)$ for $H=H'$, where $\beta$ is an inverse temperature and $\epsilon$ is the depolarising noise of the magic state. The relevant data from the spectrum of the qutrit operator $H$ enters in via $\epsilon_\star(\beta)$ and $\beta_\star$.  The vertical dashed line is the `Landauer-like' temperature threshold $\beta_\star$ and the diagonal dashed curve corresponds to the noise threshold $\epsilon_\star (\beta)$ for $\beta \leq \beta_\star$. The $\beta =0 $ line corresponds to the regime in which the magic protocol gives a unital channel.
    }
    \label{fig:rate_contour}
\end{figure}
\begin{proof}
To establish the distillation bounds we consider the distillation protocol that gives
\begin{equation}
	\E( \rho_S(\epsilon)^{\otimes n}) = \rho_S(\epsilon')^{\otimes m},
\end{equation}
for the magic state. We then consider how the protocol transforms the reference equilibrium state as
\begin{equation}
	\E( \tau^{\otimes n}) = \sigma^{\otimes m}.
\end{equation}
Since $\tau$ and $\sigma$ are assumed to be full rank stabilizer states they have a strictly positive Wigner distribution, while, in contrast, the input and output magic states will generally have quasi-probability distributions for their Wigner functions. For any such protocol we therefore have that
\begin{equation}
	(W_{\rho_n} (x), W_{\tau_n}(x) ) \succ  (W_{\rho'_m} (x), W_{\sigma_m}(x) ),
\end{equation}
where for simplicity we write $\rho_n \coloneqq \rho_S(\epsilon)^{\otimes n}$, $\rho'_m\coloneqq \rho_S(\epsilon')^{\otimes m}$ and $\tau_n\coloneqq\tau^{\otimes m}$ and $\sigma_m\coloneqq\sigma^{\otimes m}$. We then have that
\begin{equation}
	L_{\rho_n |\tau_n}(x) \ge L_{\rho'_m |\sigma_m}(x) \mbox{ for all } x.
\end{equation}
We must therefore compute the Lorenz curve data for $\rho_n$ relative to $\tau_n$, and compare with the Lorenz curve of the output state $\rho_m'$ relative to $\sigma_m$.

We define $W_{\rho | \tau}(\x) \coloneqq W_\rho(\x)/W_\tau(\x)$, which is always well-defined since $\tau$ is full-rank. Now we show in~\cref{app:wigner} that
\begin{equation}
W_{\rho_1\otimes \rho_2 | \tau_1 \otimes \tau_2} (\x_1 \oplus \x_2) = W_{\rho_1 | \tau_1}(\x_1)W_{\rho_2 | \tau_2}(\x_2),
\end{equation}
and also,
\begin{equation}
W_{\rho_1\otimes \rho_2} (\x_1 \oplus \x_2) = W_{\rho_1}(\x_1)W_{\rho_2}(\x_2),
\end{equation}
for any states $\rho_1, \rho_2$ and any full-rank stabilizer states $\tau_1, \tau_2$. \ddd{[Notation clashing here. Frame result outside as its own lemma -- we gotta check everything.]} \nick{Will frame in~\cref{app:wigner} and fix notation in the next commit}
Therefore we have that
\begin{align}
W_{\rho_n |\tau_n} (\x) &= \prod_{i=1}^n W_{\rho|\tau}(\x_i)\\
W_{\rho_n } (\x) &= \prod_{i=1}^n W_\rho(\x_i)
\end{align}
where $\x = \oplus_{i=1}^n \x_i$ is the phase space point for the full system in terms of those of the individual subsystems.

@@

Generally this is complex, but in order to obtain distillation bounds it is sufficient to determine the location of the first elbow $(x_0, L_0)$ of $L_{\rho_n|\tau_n}(\x)$. To do so, we compute the largest component 
\begin{equation}
	w_{\rm max} \coloneqq \max_{i,j,k} [\w(\rho_n | \tau_n)]_{ijk},
\end{equation}
and determine the indices at which this occurs.
Putting in the values we obtain
\begin{align}
	&(3\Z)^{-n}w_{\rm max} = \nonumber\\
	&\max\limits_{i,j,k}\Big\{ (-v)^{n-\alpha} u^{\alpha} e^{\beta (n-\alpha)E_s} e^{\beta ( i E_0 + j E_1 + k E_2 )} \Big\}, \label{eq:max_slope}
\end{align}
where $0 \leq i,j,k \leq n$ and $\alpha \coloneqq i+j+k \leq n$.
Now for $0 \leq \epsilon \leq 3/7$, we have $v \geq u$. Since we assume that $n$ is even, we need the sum $\alpha = i+j+k$ to be even too, so that the expression is positive.

@@

The rescaled distribution is then $\tilde{\w}\coloneqq (w_i/t_i)$, and we define $v_0 \coloneqq v/t_0$ and $u_i \coloneqq u/t_i$ for $i\ne 0$. We now need to consider the $n$--copy version of this, denoted $\tilde{\w}^{\otimes n}$, and determine how the largest component arises. A general component of $\tilde{\w}^{\otimes n}$ is given by,
\begin{equation}
\tilde{w}_{n,\alpha, i_1, \dots, i_8} = (-v_0)^{i_0} u_1^{i_1} u_2^{i_2} \cdots u_8^{i_8},
\end{equation}
with the assumption that firstly $n$ is always even, and the indices obeying the following conditions:
\begin{align}
0 &\le i_j \le n \mbox{ for each }j=0,1,\dots ,8 \nonumber \\
i_0+ i_1 &+ i_2 + \dots i_8 =n.
\end{align}
Now the largest entry must be positive, and so we require that $\alpha$ is an even number, and so we consider
\begin{equation}
\tilde{w}_{n,i_0, i_1, \dots, i_8} = v_0^{i_0} u_1^{i_1} u_2^{i_2} \cdots u_8^{i_8},
\end{equation}
over the above integers with $i_0 \in \{0,2,4, \dots, n\}$. Let $a = \max \{v_0, u_1, \dots ,u_8\}$, then clearly the maximum value attained ($\tilde{w}_{\rm max}\coloneqq \max_{i_0,i_1,\dots , i_8}(\tilde{w}_{n,i_0, i_1, \dots, i_8})$ ) is given by 
\begin{equation}
\tilde{w}_{\rm max} = a^n.
\end{equation}
If the numbers $\{v_0, u_1, u_2, \dots, u_8\}$ are distinct then this maximum value occurs at a unique $i_k=n$ for some $k \in \{0, 1, \dots, 8\}$. Let's consider this case first.
